Differentiable Neural Computers

### 2018-05-05 Presentation/Presentation.tex
deep LSTM sketch

\begin{frame}{Introduction: Motivation}

\end{frame}

\begin{frame}{Approach}
	Train a RNN to act as a \alert{controller} to interact with a memory matrix of N (arbitrary many) addresses.
	\pause
	\begin{enumerate}
	\item \alert{Content Lookup}\\
		\begin{itemize}
		\item \alert{Attention} over memory defined by weightings $W \in \mathbb{R}^N$
		\item Compare controller output with memory objects (\alert{auto-associative memory})
		\item Allow partial matches  (\alert{pattern completion})
		\end{itemize}
	\pause
	\item \alert{Sequential Retrieval}
		\begin{itemize} 
		\item Fill $L \in \{0,1\}^{2N}$ indexing \alert{temporal transitions}
		\item \alert{Shift} operations defined by $LW,\ L^TW$
		\end{itemize}
	\pause
	\item \alert{Dynamic Allocation}
		\begin{itemize}
		\item Mark memory locations with $\{0,1\}$ to \alert{signal usage}
		\item Manipulate signals during R/W operations to enable \alert{reallocation}
		\item Generalization to \alert{unbounded memory}
		\end{itemize}
	\end{enumerate}
\end{frame}

\begin{frame}{Controller}
	A deep \alert{long short-term memory network} receiving input:
	\[
	\xx_t=[ \pmb{x}_t; \pmb{r}_{t-1}^1; \dots ;\pmb{r}_{t-1}^R]
	\]
	and producing output:
	\[
	(\pmb{v}_t,\pmb{\xi}_t)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta)
	\]
	where $\mathcal{N}$ a set of state equations and $\vartheta$ their trainable parameters.
\end{frame}

\begin{frame}{Controller: State Equations}
	A more detailed look into $\mathcal{N}$:\\
	\begin{align}
	\ii_t^l &= \sigma(W_i^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_i^l)   \tag{\alert{input gate}}\\
	\ff_t^l &= \sigma(W_f^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_f^l)	\tag{\alert{forget gate}}\\
	\Ss_t^l &= \ff_t^l\Ss_{t-1}^l + \ii_t^ltanh(W_s^l[\xx_t;\hh_{t_1}^l;\hh_t^{l-1}] + \bb_s^l)		\tag{\alert{state}}\\
	\oo_t^l &= \sigma(W_o^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_o^l)		\tag{\alert{output gate}}\\
	\hh_t^l &= \oo_t^ltanh(\Ss_t^l)		\tag{\alert{hidden}}\\
	\nonumber \\ 
	\pmb{v}_t &= W_y[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{output vector}}\\
	\pmb{\xi}_t &= W_{\xi}[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{interface vector}}
	\end{align}
\end{frame}

@@

\begin{frame}{Controller: Outputs}
	\topskip0pt
	\vspace*{\fill}
	\[
	(\pmb{v}_t,\pmb{\xi}_t)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta)
	\]
	\pause
	Output given by:
	\[
	\pmb{y}_t = \pmb{v}_t + W_R[\pmb{r}_t^1;\dots;\pmb{r}_t^R] \tag{hello}
	\]
\end{frame}

\begin{frame}{Controller: Outputs}
	\topskip0pt
	\[
	(\pmb{v}_t,\pmb{\xi}_t)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta)
	\]
	\vspace{5pt}
	\pause
	\begin{equation*}
	\begin{split}
	\pmb\xi_t = [ \pmb{k}_t^{r,1};\dots;\pmb{k}_t^{r,R};\ 
				\hat{\beta}_t^{r,1};\dots;\hat{\beta}_t^{r,R};\ 
				\pmb{k}_t^w;\ \hat{\beta}_t^w; \\
				\hat{\pmb{e}}_t;\
				\pmb{v}_t;\ \hat{\pmb{f}}_t^1;\dots;\hat{\pmb{f}_t^R};\ 
				\hat{g}_t^a;\ \hat{g}_t^w;\ \hat{\pmb{\pi}}_t^1;\dots;\hat{\pmb{\pi}}_t^R;]					
	\end{split}
	\end{equation*}
\end{frame}

\begin{frame}{Controller: Outputs}
	\topskip0pt
	\[
	(\pmb{v}_t,\pmb{\xi}_t)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta)
	\]
	\vspace{5pt}
	\begin{equation*}
	\begin{split}
	\pmb\xi_t = &[\overbrace{\pmb{k}_t^{r,1};\dots;\pmb{k}_t^{r,R};}^\text{\tiny\alert{read keys}}\
				\overbrace{\hat{\beta}_t^{r,1};\dots;\hat{\beta}_t^{r,R};}^\text{\tiny\alert{read strengths}}\
				\overbrace{\pmb{k}_t^w;\ \hat{\beta}_t^w;}^\text{\tiny\alert{write key/strength}} \
				\overbrace{\hat{\pmb{e}}_t;}_\text{\tiny\alert{erase vector}} \\
				&\underbrace{\pmb{v}_t;}_\text{\tiny\alert{write vector}} \
				\underbrace{\hat{\pmb{f}}_t^1;\dots;\hat{\pmb{f}_t^R};}_\text{\tiny\alert{free gates}} \
				\underbrace{\hat{g}_t^a;\ \hat{g}_t^w;}_\text{\tiny\alert{allocation/write gates}}\
				\underbrace{\hat{\pmb{\pi}}_t^1;\dots;\hat{\pmb{\pi}}_t^R;}_\text{\tiny\alert{read modes}}]
	\end{split}
	\end{equation*}
\end{frame}

>>>>

\begin{frame}{Approach}
	Train a RNN to act as a \alert{controller} to interact with a memory matrix M of N (arbitrary many) addresses.
	\pause
	\begin{enumerate}
	\item \alert{Content Lookup}\\
		\begin{itemize}
		\item \alert{Attention} over memory defined by weightings $W \in \mathbb{R}^N$
		\item Compare controller output with memory objects (\alert{auto-associative memory})
		\item Allow partial matches  (\alert{pattern completion})
		\end{itemize}
	\pause
	\item \alert{Sequential Retrieval}
		\begin{itemize} 
		\item Fill $L \in \{0,1\}^{2N}$ indexing \alert{temporal transitions}
		\item \alert{Shift} operations defined by $LW,\ L^TW$
		\end{itemize}
	\pause
	\item \alert{Dynamic Allocation}
		\begin{itemize}
		\item Mark memory locations with $\{0,1\}$ to \alert{signal usage}
		\item Manipulate signals during R/W operations to enable \alert{reallocation}
		\item Generalization to \alert{unbounded memory}
		\end{itemize}
	\end{enumerate}
\end{frame}

\begin{frame}{Controller}
	A deep \alert{long short-term memory network} receiving input:
	\[
	\xx_t=[ \pmb{x}_t; \pmb{r}_{t-1}^1; \dots ;\pmb{r}_{t-1}^R]
	\]
	and producing output:
	\[
	(\pmb{\upsilon}_t,\pmb{\xi}_t)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta)
	\]
	where $\mathcal{N}$ a set of state equations and $\vartheta$ their trainable parameters.
\end{frame}

\begin{frame}{Controller: State Equations}
	A more detailed look into $\mathcal{N}$:\\
	\begin{align}
	\ii_t^l &= \sigma(W_i^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_i^l)   \tag{\alert{input gate}}\\
	\ff_t^l &= \sigma(W_f^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_f^l)	\tag{\alert{forget gate}}\\
	\Ss_t^l &= \ff_t^l\Ss_{t-1}^l + \ii_t^ltanh(W_s^l[\xx_t;\hh_{t_1}^l;\hh_t^{l-1}] + \bb_s^l)		\tag{\alert{state}}\\
	\oo_t^l &= \sigma(W_o^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_o^l)		\tag{\alert{output gate}}\\
	\hh_t^l &= \oo_t^ltanh(\Ss_t^l)		\tag{\alert{hidden}}\\
	\nonumber \\ 
	\pmb{\upsilon}_t &= W_y[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{output vector}}\\
	\pmb{\xi}_t &= W_{\xi}[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{interface vector}}
	\end{align}
\end{frame}

@@

\begin{frame}{Controller: Full View}
	\alert{LSTM Network} (multiple layers)
	\begin{figure}
	\begin{tikzpicture}[scale=1]
	\tikzset{miniLSTM/.style={draw=black, fill=white, minimum width=20pt, minimum height=10pt, thick}}
	\tikzset{split/.style={circle, draw=black,inner sep=0pt, minimum size = 3mm, fill=white}}
		\node (dummy_node) at (0,0) {};
		\foreach \name/\x in {x1/2, x2/4, x3/6, xd/8, xL/10} {\node[inner sep=0] (\name) at (\x,0) {};}
		\foreach \name/\x in {h1/3, h2/5, h3/7, hL/10} {\node[inner sep=0] (\name) at (\x, -2) {};}
		\foreach \name/\x in {s0/3, s1/5, s2/7, sL/10} {\node[split] (\name) at (\x,-3) {$\textcolor{black}{\tiny;}$};}
		\foreach \name/\x/\tag in {L1/2/$LSTM_1$, L2/4/$LSTM_2$, L3/6/$LSTM_3$, LL/10/$LSTM_L$} 
		{\node[miniLSTM] (\name) at (\x,-2) {\textcolor{black}{\tiny\tag}};}
		\node (Ld) at (8,-2) {$\dots$};
		\node (sd) at (8,-3) {$\dots$};
		\draw (dummy_node) -| (LL);
		\draw (L1) edge[->] (L2);
		\draw (L2) edge[->] (L3);
		\draw (L3) edge[->] (Ld);
		\draw (Ld) edge[->] (LL);
		\draw[->] (x1) edge (L1.north);
		\draw[->] (x2) edge (L2.north);
		\draw[->] (x3) edge (L3.north);
		\draw[->] (xL) edge (LL.north);
		\draw[->] (xd) edge (Ld.north);
		\draw[->] (1,0) |- (s0);
		\draw[->] (h1) edge (s0);
		\draw[->] (s0) edge (s1);
		\draw[->] (h2) edge (s1);
		\draw[->] (h3) edge (s2);
		\draw[->] (s1) edge (s2);
		\draw[->] (s2) edge (sd);
		\draw[->] (sd) edge (sL);
		\draw[->] (LL) edge (sL);		
		\draw[->] (sL) edge (10, -5);
		% labels
		\node[label=above right :$\scriptsize\textcolor{white}{\xx_t}$] (xt) at (0,0) {};
		\node[label=above :$\scriptsize\textcolor{white}{\hh_t^1}$] (xt) at (3,-2) {};
		\node[label=above :$\scriptsize\textcolor{white}{\hh_t^2}$] (xt) at (5,-2) {};
		\node[label=above :$\scriptsize\textcolor{white}{\hh_t^3}$] (xt) at (7,-2) {};
		\node[label=below :$\scriptsize\textcolor{white}{[\xx_t;\hh_t^1]}$] (xt) at (4,-3) {};
		\node[label=below :$\scriptsize\textcolor{white}{[\xx_t;\hh_t^1;\hh_t^2]}$] (xt) at (6,-3) {};
		\node[label=left :$\scriptsize\textcolor{white}{[\xx_t;\hh_t^1;\dots;\hh_t^L]}$] (xt) at (10,-5) {};
	\end{tikzpicture}
	\end{figure}
\end{frame}

\begin{frame}{Controller: Outputs}
	\topskip0pt
	\[
	(\pmb{\upsilon}_t,\pmb{\xi}_t)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta)
	\]
	\vspace{5pt}

	\pause
	\alert{Intermediate output} $\pmb{\upsilon}_t$
	\[
	\pmb{y}_t = \pmb{\upsilon}_t + W_R[\pmb{r}_t^1;\dots;\pmb{r}_t^R] \tag{\alert{Memory-conditioning}}
	\]

	\pause
	\alert{Interface vector} $\pmb\xi_t$

	\begin{minipage}{0.49\textwidth}
	\begin{itemize}
		\item Read keys %$\pmb{k}_t^{r,i} \in \mathbb{R}^W$
		\item Read strengths %$\beta_t^{r,i} \in [1,\infty)$
		\item Write key %$\pmb{k}_t^w \in \mathbb{R}^W$	
		\item Write strength %$\beta_t^{w} \in [1,\infty)$
		\item Erase vector %$\pmb{e}_t \in [0,1]^W$	
	\end{itemize}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
	\begin{itemize}
		\item Write vector %$\pmb{v}_t \in \mathbb{R}^W$
		\item Free gates %$\pmb{f}_t^i \in [0,1]$		
		\item Allocation gate %$g_t^a \in [0,1]$
		\item Write gate %$g_t^w \in [0,1]$
		\item Read modes %$\pmb{\pi}_t^i \in \mathcal{S}_3$
	\end{itemize}
	\end{minipage}
%	\begin{equation*}
%	\begin{split}
%	\pmb\xi_t = &[\overbrace{\pmb{k}_t^{r,1};\dots;\pmb{k}_t^{r,R};}^\text{\tiny\alert{read keys}}\
%				\overbrace{\hat{\beta}_t^{r,1};\dots;\hat{\beta}_t^{r,R};}^\text{\tiny\alert{read strengths}}\
%				\overbrace{\pmb{k}_t^w;\ \hat{\beta}_t^w;}^\text{\tiny\alert{write key/strength}} \
%				\overbrace{\hat{\pmb{e}}_t;}^\text{\tiny\alert{erase vector}} \\
%				&\underbrace{\pmb{v}_t;}_\text{\tiny\alert{write vector}} \
%				\underbrace{\hat{\pmb{f}}_t^1;\dots;\hat{\pmb{f}_t^R};}_\text{\tiny\alert{free gates}} \
%				\underbrace{\hat{g}_t^a;\ \hat{g}_t^w;}_\text{\tiny\alert{allocation/write gates}}\
%				\underbrace{\hat{\pmb{\pi}}_t^1;\dots;\hat{\pmb{\pi}}_t^R}_\text{\tiny\alert{read modes}}]
%	\end{split}
%	\end{equation*}
\end{frame}

\begin{frame}{Memory Adressing: RW}
\alert{Read operations}
\[
\pmb{r}_t^i = M_t^T\pmb{w}_t^{r,i}
\]
\pause
\alert{Write operations}
\[
M_t = M_{t-1} \circ (\pmb{1} - \pmb{w}_t^T\pmb{e}_t^T) + \pmb{w}_t^W\pmb{v}_t^T
\]
\end{frame}
